RCloud: Integrating Exploratory Visualization, Analysis and Deployment

### 2014-03-28 discussion.tex
yes

We did not anticipate having to back off so soon on the
assumption that programming is being done mainly in one
language (R). In retrospect, it should have been obvious
that many data science problems are already being solved
with cross-language approaches.

A disadvantage is that some important aspects of the environment
are more difficult to manage in RCloud than in conventional systems.
Not only does it lack an explicit separation of the development and
deployment environments, but every version of every prototype is
shared by default. While this encourages collaboration on work
in progress, errors such as misconfiguration, mismatches between
versions of packages, and ordinary programming errors are also
easily shared and can affect production websites unintentionally.
This cannot be avoided entirely, but the power of sharing is worth
this price (as proven true with distributed systems in general).
Still, this problem needs more attention if the environment is
going to be scaled up to a wider number of users with fewer
informal channels to coordinate work. Effective sharing in RCloud
will require control over versioning.
Versioning must be managed in several places: in scripts, in the R environment
such as the installed R libraries and packages, and in the external
environment such as the operating system and even protocols spoken by
remote services. At one end of the scale, we have full control over the
versioning of scripts via git, though there is still a need for conventions
for example to name stable or working versions of scripts or versions known
to work together. The R environment is under the control of its package
manager. It is at least possible for RCloud to access information about
package versions and configurations, but support for checking compatibility
and maintaining multiple versions in the same environment is not strong.
At the other end of the scale, there is not much reason to expect version
control for the external environment. Most programmers rely on clean living
and a conservative approach to upgrades.

@@

Discussion about binding times, running vs.caching.

Transparent vs. explicit operations done by programmers

Our environment assumes a high degree of trust between users.
Access control for information security is delegated to the host
operating system and web server. In practice, most operating
systems and web servers rely either on coarse permission models,
which tend to be ineffective, or on detailed access control
lists, which tend to be cumbersome and complicated. 
Information security in RCloud should be re-examined, but more
sophisticated approaches such as information flow analysis are
still an active research area, and many subtle problems remain
\cite{Moore:2011:SAF}. We are hopeful that having a richer
environment for collaboration and information sharing will encourage
practical applications for new approaches to information security
in visual analytics.

>>>>

{\bf Multiple language support.}
We did not anticipate having to back off so soon on the
assumption that programming is being done mainly in one
language (R). In retrospect, it should have been obvious
that many data science problems are already being solved
with cross-language approaches.

{\bf Versioning.}
A disadvantage is that some important aspects of the environment
are more difficult to manage in RCloud than in conventional systems.
Not only does it lack an explicit separation of the development and
deployment environments, but every version of every prototype is
shared by default. While this encourages collaboration on work
in progress, errors such as misconfiguration, mismatches between
versions of packages, and ordinary programming errors are also
easily shared and can affect production websites unintentionally.
This cannot be avoided entirely, but the power of sharing is worth
this price (as proven true with distributed systems in general).
Still, this problem needs more attention if the environment is
going to be scaled up to a wider number of users with fewer
informal channels to coordinate work. Effective sharing in RCloud
will require control over versioning.
Versioning must be managed in several places: in scripts, in the R environment
such as the installed R libraries and packages, and in the external
environment such as the operating system and even protocols spoken by
remote services. At one end of the scale, we have full control over the
versioning of scripts via git, though there is still a need for conventions
for example to name stable or working versions of scripts or versions known
to work together. The R environment is under the control of its package
manager. It is at least possible for RCloud to access information about
package versions and configurations, but support for checking compatibility
and maintaining multiple versions in the same environment is not strong.
At the other end of the scale, there is not much reason to expect version
control for the external environment. Most programmers rely on clean living
and a conservative approach to upgrades.

@@

{\bf Caching.}
An important consideration is how and where to introduce caching
in RCloud's distributed computation model. Caching can dramatically
improve performance in a way that is otherwise transparent to
application program semantics \cite{Callahan:2006:VVM, Guo:2010:TPI}.
Though caching is usually implicit, in some environments, such as
Vistrails, programmers may also have some explicit control over cache
management. This may be desirable to ensure the repeatability of
computations that rely on volatile or unreliable data sources.
For example, the stock ticker use case is one for which the data
might temporarily become unavailable, so caching could improve
reliability and consistency of results.  Alternatively, in a use case
involving text analysis through a topic modeling technique such as
Latent Dirichlet Allocation (LDA) that is relatively expensive,
we might want caching to improve performance. Currently we would
program this in RCloud by explicitly saving the analysis in a
persistent database, but it would be desirable to implement this
in a more general purpose associative cache. We envision cache
management being implemented in a new middle layer not in RCloud's
architecture today.

{\bf Security.}
% We sprinkle the security dust on at the end!
Our environment assumes a high degree of trust between users.
Access control for information security is delegated to the host
operating system and web server. In practice, most operating
systems and web servers rely either on coarse permission models,
which tend to be ineffective, or on detailed access control
lists, which tend to be cumbersome and complicated. 
Information security in RCloud should be re-examined, but more
sophisticated approaches such as information flow analysis are
still an active research area, and many subtle problems remain
\cite{Moore:2011:SAF}. We are hopeful that having a richer
environment for collaboration and information sharing will encourage
practical applications for new approaches to information security
in visual analytics.
